---
title: "projekt"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


The appendix contains data on starting salaries and personal information of high school graduates and above for each simulation study. Please use this data for the following analysis and provide a text explanation and report on the findings based on the statistical results below. The report should be approximately 3-5 pages long, including text and charts.

1 Draw a histogram of salaries (choose your own bin width).

2 Provide descriptive statistics for salaries, including sample mean, standard deviation, minimum, Q1, median, Q3, and maximum.

3 Calculate correlation coefficients between salaries and personal information (years of education, number of siblings, gender). Assign a numeric value of 1 for males and 0 for females.

4 Divide years of education into 18 years or more and less than 18 years. Try to check whether salary is related to the number of years of education.

5 Try to check if salary is related to gender.

(Hint: For questions 3 and 4, perform a two-sample t-test)


```{r}
library(readxl)
library(tidyverse)
library(ggplot2)
data <- read_xlsx("C:\\Users\\Asus\\Downloads\\data1_5.xlsx")
```

```{r}
Salary <- data$Salary
Sex <- data$Sex
Age <- data$Age
Family <- data$Family
```

```{r}
dan <- data.frame(Salary, Sex)
ggplot(dan, aes(x = Salary,  fill = Sex))+
geom_histogram(color = "black", position = "identity", bins = 20)+
  labs(title = "Histogram of salary", x = "", y = "") ## ????
```

```{r}
summary_stats <- summary(Salary)
mean_salary <- mean(Salary)
sd_salary <- sd(Salary)
min_salary <- min(Salary)
q1_salary <- quantile(Salary, 0.25)
median_salary <- median(Salary)
q3_salary <- quantile(Salary, 0.75)
max_salary <- max(Salary)
print("Statistic:")
print(summary_stats)
cat("\n")
print(paste("Sample mean: ", mean_salary))
print(paste("Standard deviation: ", sd_salary))
print(paste("Minimum value: ", min_salary))
print(paste("Q1: ", q1_salary))
print(paste("Median: ", median_salary))
print(paste("Q3: ", q3_salary))
print(paste("Maximum value: ", max_salary))
```

```{r}
data$Gender <- ifelse(data$Sex == "M", 1, 0)
cor_matrix <- cor(data[, c("Age", "Family", "Salary", "Gender")])
print(cor_matrix)
```

```{r}
less_18_sal <- data$Salary[data$Age < 18]
more_18_sal <- data$Salary[data$Age > 18]
equal_18_sal <- data$Salary[data$Age == 18]

cor_less_18 <- cor(data$Age[data$Age < 18], less_18_sal)
cor_more_18 <- cor(data$Age[data$Age > 18], more_18_sal)
cor_equal_18 <- cor(data$Age[data$Age == 18], equal_18_sal)

cat("For less then 18:", cor_less_18, "\n")
cat("For more then 18", cor_more_18, "\n")
cat("For equal 18:", cor_equal_18, "\n")  ## ????
```

```{r}
Salary_m <- data$Salary[1:10]
Salary_f <- data$Salary[11:20]
t_test_result_salary <- t.test(Salary_m, Salary_f)
print(t_test_result_salary)

  dan3 <- data.frame(Sum_salary = c(Salary_m, Salary_f), 
                     Sex = factor(rep(c("Male", "Female"), each = 10)))
  boxplot(Sum_salary ~ Sex, dan3, col = c("pink", "blue"))
  legend("topright", legend = levels(dan3$Sex), fill = c("pink", "blue"))
```

